Lesson 5
========================================================

### Third Qualitative Variable
Notes:

```{r Third Qualitative Variable}
library(ggplot2)
pf = read.csv('pseudo_facebook.tsv', sep = '\t')

ggplot(aes(x = gender, y = age),
       data = subset(pf, !is.na(gender))) + geom_boxplot() +
  stat_summary(fun.y = mean, geom = 'point', shape = 4)

```

***

### Plotting Conditional Summaries
Plot friend count for ages grouped by gender.  Then create a new dataframe that contains the same information.

```{r Plotting Conditional Summaries}
ggplot(aes(x=age, y=friend_count),
       data=subset(pf, !is.na(gender))) + 
  geom_line(aes(color = gender), stat = 'summary', fun.y = median)

library(dplyr)

pf.fc_by_age_gender = pf %>%
  filter(!is.na(gender)) %>%
  group_by(age, gender) %>%
  summarise(mean_friend_count = mean(friend_count),
            median_friend_count = median(as.numeric(friend_count)),
            n = n())

head(pf.fc_by_age_gender)

ggplot(aes(x=age, y=median_friend_count), 
           data = pf.fc_by_age_gender) +
  geom_line(aes(color = gender))
```

***

### Wide and Long Format
Long data has few columns some of which have only a couple of levels.

Wide data has fewer rows, but more columns because each column with a binary set of levels in long format is represented by two columns in the wide format.
***

### Reshaping Data
Notes:

```{r}
library(reshape2)

pf.fc_by_age_gender.wide = dcast(pf.fc_by_age_gender, 
                                 age ~ gender,
                                 value.var = "median_friend_count")

head(pf.fc_by_age_gender.wide)
```


***

### Ratio Plot
Plot the ratio of female to male friend counts.

```{r Ratio Plot}
ggplot(aes(x=age, y=female/male), 
           data = pf.fc_by_age_gender.wide) +
  geom_line() +
  geom_hline(yintercept=1, alpha=0.3, linetype=2)
```
This plot show us that young females have twice as many friends as males and the for most ages, females have more friends than males.
***

### Third Quantitative Variable
We add length of tenure of use as a third variable.  To do that, we add a new variable `year_joined` to the dataframe.

```{r Third Quantitative Variable}
pf$year_joined = 2014 - ceiling(pf$tenure /365)
summary(pf$year_joined)
table(pf$year_joined)
```

***

### Cut a Variable
Notes:
Create bins of users based on year joined.  We are turning our numeric variable to a categorical variable. This also works with continuous variables.
```{r Cut a Variable}
pf$year_joined.bucket <- cut(pf$year_joined, 
                          c(2004, 2009, 2011, 2012, 2014))
table(pf$year_joined.bucket, useNA = 'ifany')
```

***

### Plotting it All Together
Now plot the new categorical variable, year_joined.bucket.

```{r Plotting it All Together}
ggplot(aes(x=age, y=friend_count),
           data=subset(pf, !is.na(year_joined.bucket))) +
  geom_line(aes(color = year_joined.bucket), 
            stat='summary', fun.y = median)
```
This plot shows that users with longer tenures generally mave higher friend counts.
***

### Plot the Grand Mean
Add the mean across all users regardless of tenure.

```{r Plot the Grand Mean}
ggplot(aes(x=age, y=friend_count),
           data=subset(pf, !is.na(year_joined.bucket))) +
  geom_line(aes(color = year_joined.bucket), 
            stat='summary', fun.y = mean) +
  geom_line(stat='summary', fun.y = mean, linetype=2)
```

***

### Friending Rate
Create a summary for how many friends a use has for each day of tenure.

```{r Friending Rate}
with(subset(pf, tenure >= 1), summary(friend_count / tenure))
```

***
Plot the rate of friendships initiated and color by tenure buckets.
```{r Friendships Initiated}
ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket))
```

***

### Bias-Variance Tradeoff Revisited
A lot more on [Bias-Variance](http://scott.fortmann-roe.com/docs/BiasVariance.html).

```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

```

***

### Introducing the Yogurt Data Set
An article on [Scanner Data and Consumer Demand](http://www.nuff.ox.ac.uk/Economics/papers/2013/Panel%20Data%20Demand%20-%20June%2020,%202013.pdf)

***

### Histograms Revisited
Notes:

```{r Histograms Revisited}
yo = read.csv('yogurt.csv')
str(yo)

# Change the id from an int to a factor
yo$id = factor(yo$id)
str(yo)

ggplot(aes(x = price), data = yo) + 
  geom_histogram(color = 'black', fill = '#F79420') 
```
***

### Number of Purchases
The price data seems to be fairly discrete.

```{r Number of Purchases}
summary(yo)
length(unique(yo$price))
table(yo$price)

yo = transform(yo, all.purchases = strawberry + blueberry + pina.colada + plain + mixed.berry)

qplot(x=all.purchases, data = yo, binwidth = 1,
      fill=I('#099DD9'))
```

***

### Prices over Time
Notes:

```{r Prices over Time}
ggplot(aes(x=time, y=price), data = yo) +
  geom_jitter(alpha=1/4, shape=21, fill=I('#F79420'))
```

***

### Sampling Observations
When using data that has multiple observations over time, it is useful to sample some of the observations.  This can help to determine within goup and between group variability.

***

### Looking at Samples of Households
This plot shows the price paid of certain households over time.  The size of the points show number of purchases on each occasion.
```{r Looking at Sample of Households}
set.seed(4230)
sample.ids = sample(levels(yo$id), 16)

ggplot(aes(x=time, y=price),
       data = subset(yo, id %in% sample.ids)) +
  facet_wrap( ~ id) +
  geom_line() +
  geom_point(aes(size = all.purchases), pch=1)

```

***

### Scatterplot Matrix
Creates a grid of scatterplots between each of the variables.

```{r}
library(GGally)
theme_set(theme_minimal(20))

set.seed(1836)
pf_subset = pf[, c(2:15)]
names(pf_subset)
ggpairs(pf_subset[sample.int(nrow(pf_subset), 1000),])
```

***

### Heat Maps
Notes:

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
library(reshape2)
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```